A Large-scale Study of Polymorphism in R

### 2019-03-26 paper/main.tex
updated counts and new signatures, dunno if i should push the data (it big)

% https://www.r-pkg.org/downloaded
\begin{figure}[!th]{\footnotesize\begin{tabular}{@{}r||l|r|r|r|r|r@{}}\hline
\bf Package & \bf Description & \bf R LOC &\bf C LOC &\bf Scripts & \bf Calls Observed & \bf Calls Recorded \\
\hline
\tt Rcpp  & Seamless C++ integration & 2.1K & 4K & 25 & 51K & 282 \\
\tt rlang & Functions for 'Tidyverse' & 4.4K & 3.8K & 122 & 2,240K & 5,171 \\
\tt glue  & Interpreted string literals & 0.2K & 0.2K & 8 & 2K & 120 \\
\tt tibble & Simple data frames & 1.3K & 0.3K & 16 & 839K & 4,148 \\
\tt stringi &  String processing & 1.5K & 515K & 64 & 692K & 666 \\
\tt ggplot2 & Data visualisations & 10.8K & 0 & 130 & 110K & 2,636\\
\tt dplyr  &  Data manipulation & 3.8K & 4.7K & 78 & 157K & 1,967 \\
\tt pillar & Formatting for columns & 1.3K & 0 & 13 & 549K & 1,273 \\
\tt R6 & Classes w. ref. semantics & 0.6K & 0 & 2 & 1K & 226 \\
\tt stringr & String operations & 0.5K & 0 & 32 & 1,551K & 371 \\
\end{tabular}}\caption{10 Most Downloaded Packages.}\label{most}
\end{figure}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsFGTS}
\centering
\begin{tabular}{rlrr}
  \hline
 & Argument Polymorphism & Count & Percentage \\ 
  \hline
1 & Full Monomorphic & 457K & 87.75 \\ 
  2 & Monomorphic in Type & 485K & 93.08 \\ 
  3 & Monomorphic in Class & 480K & 92.16 \\ 
  4 & Monomorphic in Attribute Pattern & 480K & 91.99 \\ 
  5 & Total Seen & 521K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the FGTS.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsFGTS}
\centering
\begin{tabular}{rlrr}
  \hline
 & Function Polymorphism & Count & Percentage \\ 
  \hline
1 & Full Monomorphic & 86K & 69.44 \\ 
  2 & Monomorphic in Type & 100K & 81.01 \\ 
  3 & Monomorphic in Class & 97K & 78.45 \\ 
  4 & Monomorphic in Attribute Pattern & 96K & 77.47 \\ 
  5 & Total Seen & 124K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the FGTS.}
\end{table}

@@

In the tables, ``Full Monomorphic'' denotes monomorphism in type, class, and attribute.
We see that roughly 7\% of all argument positions in R are polymorphic in type.
The numbers for attribute and class polymorphism are included for
completeness, but the important number is that of type polymorphism (as our
method of constructing types sometimes factors in class and attribute).
Specifically for classes, often the class and \code{typeof} a value will
line up (refer to Section~\ref{sec:rlang}).

As for functions, roughly 20\% are polymorphic in type, or put differently 20\% have at least one polymorphic argument or a polymorphic return.
The discrepancy between this and the polymorphicity of arguments is explained by the tendency of some functions to have {\it many} arguments.

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 21:27:37 2019
\begin{table}[ht]
\label{tab:toppolyFDTS}
\centering
\begin{tabular}{lrr}
  \hline
 Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
  vector/double, vector/integer & 4838 & 13.42 \\ 
  scalar/double, scalar/integer & 3666 & 10.17 \\ 
  NULL, vector/double & 1463 & 4.06 \\ 
  data.frame, vector/double & 1324 & 3.67 \\ 
  NULL, scalar/character & 1237 & 3.43 \\ 
   NULL, scalar/double & 1174 & 3.26 \\ 
   list$<$any$>$, list$<$double$>$ & 1092 & 3.03 \\ 
   list$<$any$>$, list$<$list$>$ & 767 & 2.13 \\ 
   scalar/character, scalar/double & 674 & 1.87 \\ 
   NULL, vector/character & 640 & 1.78 \\ 
     \hline
\end{tabular}
\caption{Top polymorphic argument signatures in the FDTS.}
\end{table}

>>>>

% https://www.r-pkg.org/downloaded
\begin{figure}[!th]{\footnotesize\begin{tabular}{@{}r||l|r|r|r|r|r@{}}\hline
\bf Package & \bf Description & \bf R LOC &\bf C LOC &\bf Scripts & \bf Calls Observed & \bf Calls Recorded \\
\hline
\tt Rcpp  & Seamless C++ integration & 2.1K & 4K & 25 & 55K & 282 \\
\tt rlang & Functions for 'Tidyverse' & 4.4K & 3.8K & 122 & 3,924K & 5,171 \\
\tt glue  & Interpreted string literals & 0.2K & 0.2K & 8 & 4K & 120 \\
\tt tibble & Simple data frames & 1.3K & 0.3K & 16 & 1,332K & 4,148 \\
\tt stringi &  String processing & 1.5K & 515K & 64 & 923K & 666 \\
\tt ggplot2 & Data visualisations & 10.8K & 0 & 130 & 153K & 2,636\\
\tt dplyr  &  Data manipulation & 3.8K & 4.7K & 78 & 233K & 1,967 \\
\tt pillar & Formatting for columns & 1.3K & 0 & 13 & 803K & 1,273 \\
\tt R6 & Classes w. ref. semantics & 0.6K & 0 & 2 & 1K & 226 \\
\tt stringr & String operations & 0.5K & 0 & 32 & 1,764K & 371 \\
\end{tabular}}\caption{10 Most Downloaded Packages.}\label{most}
\end{figure}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsFGTS}
\centering
\begin{tabular}{rlrr}
  \hline
 & Argument Polymorphism & Count & Percentage \\ 
  \hline
  1 & Full Monomorphic & 784K & 85.19 \\ 
  2 & Monomorphic in Type & 839K & 91.27 \\ 
  3 & Monomorphic in Class & 831K & 90.31 \\ 
  4 & Monomorphic in Attribute Pattern & 830K & 90.25 \\ 
  5 & Total Seen & 920K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the FGTS.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsFGTS}
\centering
\begin{tabular}{rlrr}
  \hline
 & Function Polymorphism & Count & Percentage \\ 
  \hline
  1 & Full Monomorphic & 143K & 65.64 \\ 
  2 & Monomorphic in Type & 169K & 77.65 \\ 
  3 & Monomorphic in Class & 163K & 75.13 \\ 
  4 & Monomorphic in Attribute Pattern & 161K & 74.31 \\ 
  5 & Total Seen & 217K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the FGTS.}
\end{table}

@@

In the tables, ``Full Monomorphic'' denotes monomorphism in type, class, and attribute.
We see that roughly 9\% of all argument positions in R are polymorphic in type.
The numbers for attribute and class polymorphism are included for
completeness, but the important number is that of type polymorphism (as our
method of constructing types sometimes factors in class and attribute).
Specifically for classes, often the class and \code{typeof} a value will
line up (refer to Section~\ref{sec:rlang}).

As for functions, roughly 22\% are polymorphic in type, or put differently 22\% have at least one polymorphic argument or a polymorphic return.
The discrepancy between this and the polymorphicity of arguments is explained by the tendency of some functions to have {\it many} arguments.

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 21:27:37 2019
\begin{table}[ht]
\label{tab:toppolyFDTS}
\centering
\begin{tabular}{lrr}
  \hline
 Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
vector/double, vector/integer & 10,316 & 12.85 \\ 
scalar/double, scalar/integer & 7,194 & 8.96 \\ 
NULL, vector/double & 3,087 & 3.85 \\ 
data.frame, vector/double & 2,604 & 3.24 \\ 
NULL, scalar/character & 2,333 & 2.91 \\ 
list$<$any$>$, list$<$double$>$ & 2,224 & 2.77 \\ 
NULL, scalar/double & 2,066 & 2.57 \\ 
raw\_NA, vector/double & 1,465 & 1.83 \\ 
list$<$any$>$, list$<$list$>$ & 1,426 & 1.78 \\ 
NULL, vector/character & 1,385 & 1.73 \\ 
     \hline
\end{tabular}
\caption{Top polymorphic argument signatures in the FDTS.}
\end{table}
